\bf Comparing the Power of Plot Designs to Reveal Correlation

### 2015-12-15 paper/paper.Rnw
few code fixes

<<setup, echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
library(lme4)
library(ggplot2)
library(mnormt)
library(tidyr)
# Convert reshape2 to tidyr, melt to gather - done
#library(reshape2)
library(nullabor)
library(dplyr)
library(mFilter)
library(grid)
library(gridExtra)
library(stringr)
#library(splitstackshape) - changed cSplit to unnest
#library(readr)
#library(lubridate)
# Convert these to read_csv at some point
correl <- read.csv("correlations.csv", stringsAsFactors = FALSE)
full <- read.csv("full.csv", row.names=1, stringsAsFactors = FALSE) 
Turk <- read.csv("turk.csv", stringsAsFactors = FALSE)

@@

\begin{figure}[htbp]
  \centering
%<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
\iffalse
# correl <- read.csv("correlations.csv") # You have already read this
pin <- select(full, pic_id, response_0, actual_location, 
              test_param, smoothed, n) #full[,c(5, 6, 10, 15, 16, 17)]
# Need to use the tidyr commands instead of cSplit
# Here's where we will need to make some changes to incorporate the multiple picks
#pin <- pin %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
#pin$response_0 <- separate(pin, response_0, "response_0", sep=",", extra="drop")
pin$choice <- separate(pin, response_0, c("choice", "l1", "l2"), ",", extra="drop")$choice
pin$choice[str_detect(pin$actual_location, pin$response_0)] <- 
  pin$actual_location[str_detect(pin$actual_location, pin$response_0)]
                         #as.integer(pin$response_0)
counts <- data.frame(table(pin$pic_id, pin$choice))
counts$Var1 <- as.numeric(as.character(counts$Var1))
counts$Var2 <- as.numeric(as.character(counts$Var2))
colnames(counts) <- c("lineup", "plot", "count")
pin <- merge(counts, Turk, by.x="lineup", by.y="pic_id")
pin <- select(pin, lineup, plot, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
pin$r <- factor(pin$r)
colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines"), smoothed=ifelse(smoothed=="no", 0, 1))
pin$base <- 0
pin$Correlation <- diag(as.matrix(correl[pin$lineup, (pin$plot+2)]))
pin <- mutate(pin, true = ifelse(plot==true_pos, "True plot", "Null plot"))
pin <- merge(pin, Turk[,c(2, 16)], by.x = "lineup", by.y = "pic_id")
pin <- mutate(pin, percent = count/attempt)
ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
       geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + labs(title="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
#Code to be fixed
\fi
%@

@@

<<powerfit, dependson='setup', echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
full$subj_id <- factor(full$ip_address, labels=1:207)

# Need to keep only evaluations done by subject who completed at least 10
x <- data.frame(table(full$subj_id))
keep <- x$Var1[x$Freq>9]
full.sub <- filter(full, subj_id %in% keep)
full.sub$r <- round(full.sub$correlation, 1)
full.sub$rsgn <- ifelse(full.sub$r >= 0, 1, 0)
full.sub$rabs <- abs(full.sub$r)
full.sub$rabs <- factor(full.sub$rabs)
full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
full.sub <- select(full.sub, subj_id, correct, r, rsgn, smoothed, n, plot, log_time)#[,c(1,7, 15,16,17,20:26)]
#full.sub$rpos <- ifelse(full.sub$r > 0, full.sub$r, 0)
#full.sub$rneg <- ifelse(full.sub$r < 0, full.sub$r, 0)
full.sub$alpha <- log(.05/0.95)

m8l <- glmer(correct~offset(alpha) + r + rsgn + r*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial")
m8s <- glmer(correct~offset(alpha) + r + rsgn + r*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "S"), family="binomial")
@
%Code to make a table to be added

@@

full.pred8s <- data.frame(expand.grid(r=r, n=n, alpha=log(.05/0.95),
                                      smoothed=c(0,1), 
                                      subj_id=subset(full.sub, plot=="S")$subj_id)) #ranef(m4)[[1]][,1]))
full.pred8s$rsgn <- ifelse(full.pred8s$r > 0, 1, 0)
full.pred8s$rabs <- abs(full.pred8s$r)
full.pred8s$fitre <- predict(m8s, full.pred8s, type="response", re.form=~(1|subj_id),
                             allow.new.levels=TRUE)
full.pred8s$fitfx <- predict(m8s, full.pred8s, type="response", re.form=NA,
                             allow.new.levels=TRUE)
full.pred8s$plot <- ifelse(full.pred8s$smoothed == 0, "Scatterplot", "Scatterplot (Smoothed)")
full.pred8s <- filter(full.pred8s, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)
full.pred2 <- rbind(full.pred8s, full.pred8l)

full.pred2 <- mutate(full.pred2, plot2 = ifelse(substr(plot, 1, 1) =="S", "Scatterplot", "Overlaid Lines"))

qplot(r, fitre, data=full.pred2, color=plot, 
      group=interaction(subj_id, plot), geom="line", alpha=I(0.1)) + 
  scale_color_manual("Plot Design", values=c("#33AA33", "#9977EE", "#FF9933", "#00DDAA")) + 
  xlab("Correlation") + ylab("Percent Detected") + labs(title="Predicted Power") +
  geom_line(aes(y=fitfx), size=2, alpha=1) + theme_bw() + facet_wrap(~n, ncol = 2) + theme(aspect.ratio=1.2)
@
  \label{powermodel}
  \caption{Predicted power fit from the mixed effects logit power model. The weakness of the overlaid line graph in negative correlation and  the scatter plot's gain from sample size are highlighted.}
\end{figure}

@@

<<timefit, dependson='powerfit', echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + plot + plot*smoothed 
                  + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
@
%get table code

>>>>

<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
library(lme4)
library(ggplot2)
library(mnormt)
library(tidyr)
# Convert reshape2 to tidyr, melt to gather - done
#library(reshape2)
library(nullabor)
library(dplyr)
library(mFilter)
library(grid)
library(gridExtra)
library(stringr)
library(pander)
#library(splitstackshape) - changed cSplit to unnest
#library(readr)
#library(lubridate)
# Convert these to read_csv at some point
correl <- read.csv("correlations.csv", stringsAsFactors = FALSE)
full <- read.csv("full.csv", row.names=1, stringsAsFactors = FALSE) 
Turk <- read.csv("turk.csv", stringsAsFactors = FALSE)

@@

\begin{figure}[htbp]
  \centering
<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
# correl <- read.csv("correlations.csv") # You have already read this
pin <- select(full, pic_id, response_0, actual_location, 
              test_param, smoothed, n) #full[,c(5, 6, 10, 15, 16, 17)]
# Need to use the tidyr commands instead of cSplit
# Here's where we will need to make some changes to incorporate the multiple picks
#pin <- pin %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
#pin$response_0 <- separate(pin, response_0, "response_0", sep=",", extra="drop")
pin$choice <- separate(pin, response_0, c("choice", "l1", "l2"), ",", extra="drop")$choice
pin$choice[str_detect(pin$actual_location, pin$response_0)] <- 
  pin$actual_location[str_detect(pin$actual_location, pin$response_0)]
                         #as.integer(pin$response_0)
counts <- data.frame(table(pin$pic_id, pin$choice))
counts$Var1 <- as.numeric(as.character(counts$Var1))
counts$Var2 <- as.numeric(as.character(counts$Var2))
colnames(counts) <- c("lineup", "plot", "count")
pin <- merge(counts, Turk, by.x="lineup", by.y="pic_id")
pin <- select(pin, lineup, plot.x, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
pin$r <- factor(pin$r)
colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines"))#, smoothed=ifelse(smoothed=="no", 0, 1))
pin$base <- 0
pin$Correlation <- diag(as.matrix(correl[pin$lineup, (pin$plot+2)]))
pin <- mutate(pin, true = ifelse(plot==true_pos, "True plot", "Null plot"))
pin <- merge(pin, Turk[,c(2, 16)], by.x = "lineup", by.y = "pic_id")
pin <- mutate(pin, percent = count/attempt)
ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
       geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + labs(title="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
#Code to be fixed
@

@@

<<powerfit, dependson='setup', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
full$subj_id <- factor(full$ip_address, labels=1:207)

# Need to keep only evaluations done by subject who completed at least 10
x <- data.frame(table(full$subj_id))
keep <- x$Var1[x$Freq>9]
full.sub <- filter(full, subj_id %in% keep)
full.sub$r <- round(full.sub$correlation, 1)
full.sub$rsgn <- ifelse(full.sub$r >= 0, 1, 0)
full.sub$rabs <- abs(full.sub$r)
#full.sub$rabs <- factor(full.sub$rabs)
full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
full.sub <- select(full.sub, subj_id, correct, r, rsgn, rabs, smoothed, n, plot, log_time, response_0, actual_location)#[,c(1,7, 15,16,17,20:26)]
full.sub$attempts <- sapply(strsplit(as.character(full.sub$response_0), ","), length)
#full.sub$weights = 1/(full.sub$attempts) #weights are actually number of attempts in the glm
#full.sub <- full.sub %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
#full.sub$correct <- ifelse(full.sub$response_0 == full.sub$actual_location, 1, 0) #fixing correct for multiple guesses
#full.sub$rpos <- ifelse(full.sub$r > 0, full.sub$r, 0)
#full.sub$rneg <- ifelse(full.sub$r < 0, full.sub$r, 0)
full.sub$alpha <- log(.05/0.95)
full.sub$correct.prop <- full.sub$correct/full.sub$attempts #correct.prop - proportion of correct selections, weights - number of attempts

m8l <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial", weights=attempts)
m8s <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "S"), family="binomial", weights=attempts)
pander(summary(m8l), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for overlaid lines"))
pander(summary(m8s), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for scatter plots"))
@

@@

full.pred8s <- data.frame(expand.grid(r=r, n=n, alpha=log(.05/0.95),
                                      smoothed=c(0,1), 
                                      subj_id=subset(full.sub, plot=="S")$subj_id)) #ranef(m4)[[1]][,1]))
full.pred8s$rsgn <- ifelse(full.pred8s$r > 0, 1, 0)
full.pred8s$rabs <- abs(full.pred8s$r)
full.pred8s$fitre <- predict(m8s, full.pred8s, type="response", re.form=~(1|subj_id),
                             allow.new.levels=TRUE)
full.pred8s$fitfx <- predict(m8s, full.pred8s, type="response", re.form=NA,
                             allow.new.levels=TRUE)
full.pred8s$plot <- ifelse(full.pred8s$smoothed == 0, "Scatterplot", "Scatterplot (Smoothed)")
full.pred8s <- filter(full.pred8s, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)
full.pred <- rbind(full.pred8s, full.pred8l)

#full.pred <- mutate(full.pred2, plot2 = ifelse(substr(plot, 1, 1) =="S", "Scatterplot", "Overlaid Lines"))

qplot(r, fitre, data=full.pred, color=plot, 
      group=interaction(subj_id, plot), geom="line", alpha=I(0.1)) + 
  scale_color_manual("Plot Design", values=c("#33AA33", "#9977EE", "#FF9933", "#00DDAA")) + 
  xlab("Correlation") + ylab("Estimated Power") + labs(title="Predicted Power") +
  geom_line(aes(y=fitfx), size=2, alpha=1) + theme_bw() + facet_wrap(~n, ncol = 2) + theme(aspect.ratio=1.2)
@
  \label{powermodel}
  \caption{Predicted power fit from the mixed effects logit power model. The weakness of the overlaid line graph in negative correlation and  the scatter plot's gain from sample size are highlighted.}
\end{figure}

@@

<<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + plot + plot*smoothed 
                  + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
pander(summary(timelmer), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for time taken"))
@
